---
title: "Final Project Visualizations"
author: "Jonathan Pedroza"
date: "February 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages(c("ggpubr"))
# devtools::install_github("hrbrmstr/ggalt")
library(tidyverse)
library(here)
library(rio)
library(nhanesA)
library(xaringan)
library(ggridges)
library(viridis)
library(ggalt)
library(ggpubr)
library(RColorBrewer)

```

```{r}
# Initial time to get data from nhanes datasets

# demographics from 2011-2012
demo_g <- nhanes('DEMO_G')

# demographics from 2013-2014
demo_h <- nhanes('DEMO_H')

# demographics from 2015-2016
demo_i <- nhanes('DEMO_I')


# physical activity data frames 2011-2012
paq_g <- nhanes('PAQ_G')

# physical activity data frames 2013-2014
paq_h <- nhanes('PAQ_H')

# physical activity data frames 2015-2016
paq_i <- nhanes('PAQ_I')


# Example of exporting data into csv files
write.csv(demo_g, "demo_g.csv")
write.csv(demo_h, "demo_h.csv")
write.csv(demo_i, "demo_i.csv")

write.csv(paq_g, "paq_g.csv")
write.csv(paq_h, "paq_h.csv")
write.csv(paq_i, "paq_i.csv")

```

```{r loading csv files}
 
demo_g <- import(here("data", "demo_g.csv"),
                setclass = "tbl_df") %>% 
  janitor::clean_names()
demo_h <- import(here("data", "demo_h.csv"),
                setclass = "tbl_df") %>% 
  janitor::clean_names()
demo_i <- import(here("data", "demo_i.csv"),
                setclass = "tbl_df") %>% 
  janitor::clean_names()

paq_g <- import(here("data", "paq_g.csv"),
                setclass = "tbl_df") %>% 
  janitor::clean_names()
paq_h <- import(here("data", "paq_h.csv"),
                setclass = "tbl_df") %>% 
  janitor::clean_names()
paq_i <- import(here("data", "paq_i.csv"),
                setclass = "tbl_df") %>% 
  janitor::clean_names()


```

```{r}
demo_g %>% 
  count(seqn) %>% 
  filter(n>1)

demo_h %>% 
  count(seqn) %>% 
  filter(n>1)

demo_i %>% 
  count(seqn) %>% 
  filter(n>1)

paq_g %>% 
  count(seqn) %>% 
  filter(n>1)

paq_h %>% 
  count(seqn) %>% 
  filter(n>1)

paq_i %>% 
  count(seqn) %>% 
  filter(n>1)
```


```{r joins}

super_demo <- full_join(demo_g, demo_h)
demo <- full_join(super_demo, demo_i)

super_pa <- full_join(paq_g, paq_h)
pa <- full_join(super_pa, paq_i)

demo %>%
  count(seqn) %>% 
  filter(n>1)

pa %>% 
  count(seqn) %>% 
  filter(n>1)

complete <- left_join(demo, pa, by = "seqn")

```

```{r}
psych::describe(complete, na.rm = TRUE)

complete <- complete %>%
  dplyr::select(seqn,
         sddsrvyr,
         riagendr,
         ridageyr,
         ridreth3,
         dmdborn4,
         dmdcitzn,
         dmdyrsus,
         dmdeduc2,
         dmdfmsiz,
         indhhin2,
         pad615,
         pad630,
         pad660, 
         pad675, 
         pad680, 
         paq706,
         paq710,
         paq715)
```


```{r selecting specific variables}

complete <- complete %>%
   naniar::replace_with_na(replace = list(pad615 = c(7777, 9999),
     paq715 = c(77, 99),
     pad630 = c(7777, 9999),
     pad660 = c(7777, 9999),
     pad675 = c(7777, 9999),
     pad680 = c(7777, 9999),
     paq706 = c(77, 99),
     paq710 = c(77, 99)))

complete <- complete %>%
  mutate(data_cycle = case_when(sddsrvyr == 7 ~ "2011_2012",
                            sddsrvyr == 8 ~ "2013_2014",
                            sddsrvyr == 9 ~ "2015_2016",
                            TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(gender = case_when(riagendr == 1 ~ "male",
                            riagendr == 2 ~ "female",
                            TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(age = ridageyr)

complete <- complete %>% 
  mutate(race = case_when(ridreth3 == 1 ~ "mexican",
                       ridreth3 == 2 ~ "other_hispanic",
                       ridreth3 == 3 ~ "white",
                       ridreth3 == 4 ~ "black",
                       ridreth3 == 6 ~ "asian",
                       ridreth3 == 7 ~ "multi",
                       TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(birth_country = case_when(dmdborn4 == 1 ~ "us_born",
                                dmdborn4 == 2 ~ "other",
                                TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(citizen = case_when(dmdcitzn == 1 ~ "citizen",
                          dmdcitzn == 2 ~ "not_citizen",
                          TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(length_us = case_when(dmdyrsus == 1 ~ "less_year",
                            dmdyrsus == 2 ~ "1_5yr",
                            dmdyrsus == 3 ~ "5_10yr",
                            dmdyrsus == 4 ~ "10_15yr",
                            dmdyrsus == 5 ~ "15_20yr",
                            dmdyrsus == 6 ~ "20_30yr",
                            dmdyrsus == 7 ~ "30_40yr",
                            dmdyrsus == 8 ~ "40_50yr",
                            dmdyrsus == 9 ~ "50_plus_yr",
                            TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(ed = case_when(dmdeduc2 == 1 ~ "less_9",
                     dmdeduc2 == 2 ~ "9_11",
                     dmdeduc2 == 3 ~ "highschool",
                     dmdeduc2 == 4 ~ "aa",
                     dmdeduc2 == 5 ~ "college_grad",
                     TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(household = case_when(dmdfmsiz == 1 ~ "1",
                            dmdfmsiz == 2 ~ "2",
                            dmdfmsiz == 3 ~ "3",
                            dmdfmsiz == 4 ~ "4",
                            dmdfmsiz == 5 ~ "5",
                            dmdfmsiz == 6 ~ "6",
                            dmdfmsiz == 7 ~ "7_plus",
                            TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(income = case_when(indhhin2 == 1 ~ "0_5k",
                         indhhin2 == 2 ~ "5_10k",
                         indhhin2 == 3 ~ "10_15k",
                         indhhin2 == 4 ~ "15_20k",
                         indhhin2 == 5 ~ "20_25k",
                         indhhin2 == 6 ~ "25_35k",
                         indhhin2 == 7 ~ "35_45k",
                         indhhin2 == 8 ~ "45_55k",
                         indhhin2 == 9 ~ "55_65k",
                         indhhin2 == 10 ~ "65_75k",
                         indhhin2 == 12 ~ NA_character_,
                         indhhin2 == 13 ~ NA_character_,
                         indhhin2 == 14 ~ "75_100k",
                         indhhin2 == 15 ~ "100_plus",
                         TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(vig_work = pad615)

complete <- complete %>% 
  mutate(mod_work = pad630)

complete <- complete %>% 
  mutate(vig_rec_pa = pad660)

complete <- complete %>% 
  mutate(mod_rec_pa = pad675)

complete <- complete %>% 
  mutate(sed = pad680)

complete <- complete %>% 
  mutate(days_active_60 = case_when(paq706 == 0 ~ "0_day",
                                 paq706 == 1 ~ "1_day",
                                 paq706 == 2 ~ "2_day",
                                 paq706 == 3 ~ "3_day",
                                 paq706 == 4 ~ "4_day",
                                 paq706 == 5 ~ "5_day",
                                 paq706 == 6 ~ "6_day",
                                 paq706 == 7 ~ "7_day",
                                 TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(hours_tv = case_when(paq710 == 0 ~ "less_hour",
                             paq710 == 1 ~ "1_hour",
                             paq710 == 2 ~ "2_hours",
                             paq710 == 3 ~ "3_hours",
                             paq710 == 4 ~ "4_hours",
                             paq710 == 5 ~ "5_hours_plus",
                             paq710 == 8 ~ "no_tv",
                           TRUE ~ NA_character_))

complete <- complete %>% 
  mutate(hours_comp = case_when(paq715 == 0 ~ "less_hour",
                             paq715 == 1 ~ "1_hour",
                             paq715 == 2 ~ "2_hours",
                             paq715 == 3 ~ "3_hours",
                             paq715 == 4 ~ "4_hours",
                             paq715 == 5 ~ "5_hours_plus",
                             paq715 == 8 ~ "no_comp",
                             TRUE ~ NA_character_))

```

```{r}

data <- complete %>%
  select(-2:-19)

data$age_cut <- cut(data$age, breaks = c(0, 18, 25, 55, Inf),
                    labels = c("minor", "young adult", "adult",
                               "older adult"))

```

## Visualizations set #1
```{r}
data %>% 
  filter(age > 0) %>% 
ggplot(aes(age, fill = age_cut)) +
  geom_histogram(color = "white",
                 alpha = .7,
                 bins = 50) + 
  facet_wrap(~data_cycle) +
  labs(x = "Age",
       y = "Count",
       Title = "Breakdown of Participants by Age",
       subtitle = "Separated by Age Group",
       caption = "NHANES data from 2011-2012, 2013-2014, 2015-2016") +
  theme_minimal() +
  guides(fill = "none")

# Part of assignment to examine differences in sedentary behaviors by race and income

p1 <- ggplot(data, aes(gender, sed)) +
  geom_col(alpha = .7, position = "dodge2", aes(fill = race)) +
  scale_y_continuous(limits = c(0, 1000)) +
  scale_fill_viridis_d() +
  facet_wrap(~data_cycle) +
  theme_minimal() +
    labs(x = "Gender",
       y = "Sedentary Behaviors",
       title = "Gender Physical Activity & Sedentary Behaviors",
       subtitle = "Separated by Different Racial/Ethnic Groups", 
       caption = " ") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(data, aes(gender, sed)) +
  geom_col(alpha = .7, position = "dodge2", aes(fill = income)) +
  scale_y_continuous(limits = c(0, 1000)) +
  scale_fill_viridis_d() +
  theme_minimal() +
  facet_wrap(~data_cycle) +
    labs(x = "Gender",
       y = "Sedentary Behaviors",
       title = "Gender Physical Activity & Sedentary Behaviors",
       subtitle = "Separated by Different Racial/Ethnic Groups",
       caption = " ") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Part of the assignment to examine differences in activity by race and income

p3 <- ggplot(data, aes(gender, mod_rec_pa)) +
  geom_col(alpha = .7, position = "dodge2", aes(fill = race)) +
  scale_y_continuous(limits = c(0, 1000)) +
  scale_fill_viridis_d() +
  theme_minimal() +
  facet_wrap(~data_cycle) +
    labs(x = "Gender",
       y = "Moderate Physical Activity",
       title = " ",
       subtitle = " ",
       caption = "NHANES data from 2011-2012, 2013-2014, 2015-2016") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p4 <- ggplot(data, aes(gender, mod_rec_pa)) +
  geom_col(alpha = .7, position = "dodge2", aes(fill = income)) +
  scale_y_continuous(limits = c(0, 1000)) +
  scale_fill_viridis_d() +
  theme_minimal() +
  facet_wrap(~data_cycle) +
    labs(x = "Gender",
       y = "Moderate Physical Activity",
       title = " ",
       subtitle = " ",
       caption = "NHANES data from 2011-2012, 2013-2014, 2015-2016") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggarrange(p1, p3,
          ncol = 2, nrow = 1)

ggarrange(p2, p4,
          ncol = 2, nrow = 1)
```

###Feedback
This set of graphs is a really good starting point for your portfolio and I thought you were strategic in the order that you set your plots up in giving the overall frequency picture in the first graph, then moving into the specifics with the subsequent ones. I used the argument ggarrange() to package the Activity and Sedentary plots together, but you might even think about displaying all four graphs together as one visualization. The use of the age cutoffs by color really helps the distribution to be visualized in chucks, so great call on that. One area that could help the Physical Activity plots is to clean up the x-axis ticks by using theme(). One other suggested change that I played with was adding theme_minimal().

## Visualization Set Two
```{r}
his_data <- data %>% 
  filter(race == "mexican" | race == "other_hispanic")

less_yr <- data %>%
  filter(race == "mexican" | race == "other_hispanic") %>% 
  filter(length_us == "less_year")

fifty_plus <- data %>% 
  filter(race == "mexican" | race == "other_hispanic") %>% 
  filter(length_us == "50_plus_yr")

# These two plots are one plot for my project
# the first is looking at working physical activity and then the next one is similar
# with recreational physical activity
p5 <- ggplot(his_data, aes(mod_work, vig_work)) +
  geom_jitter(aes(color = race)) +
  geom_encircle(aes(mod_work, vig_work),
                data = less_yr,
                color = "red",
                size = 2,
                expand = .04) +
  geom_encircle(aes(mod_work, vig_work),
                data = fifty_plus,
                color = "blue",
                size = 2,
                expand = .04) +
  facet_wrap(~data_cycle) +
  theme_minimal() +
  scale_color_viridis_d()

p6<- ggplot(his_data, aes(mod_rec_pa, vig_rec_pa)) +
  geom_jitter(aes(color = race), size = 2) +
  geom_encircle(aes(mod_rec_pa, vig_rec_pa),
                data = less_yr,
                color = "red",
                size = 2,
                expand = .04) +
  geom_encircle(aes(mod_rec_pa, vig_rec_pa),
                data = fifty_plus,
                color = "blue",
                size = 2,
                expand = .04) +
  facet_wrap(~data_cycle) +
  theme_minimal() +
  scale_color_viridis_d()

ggarrange(p5, p6,
          ncol = 1, nrow = 2)

```

###Feedback
Very cool! I had no idea that there was a ggencircle and I think it really helps draw attention to categorical differences. I went ahead and used ggarrange() again and added theme_minimal to allow the colors to really pop. 

##Visualization Set Three
```{r}
# exploratory plots

ggplot(data, aes(forcats::fct_reorder(data_cycle, sed), sed)) +
  geom_col(alpha = .7, position = "dodge2", aes(fill = race)) +
  scale_y_continuous(limits = c(0, 1250)) +
  scale_fill_viridis_d(option = "magma") +
  theme_minimal()


ggplot(data, aes(data_cycle)) +
  geom_bar(aes(fill = fct_reorder(race, age)), position = "dodge2") +
  scale_fill_brewer(palette = "Dark2")

ggplot(data, aes(race)) +
  geom_bar(position = "dodge2", aes(fill = income)) +
  facet_wrap(~data_cycle) +
  coord_flip() +
  scale_fill_viridis_d() +
  theme_minimal()

ggplot(data, aes(gender, days_active_60)) +
  geom_tile(aes(fill = age)) +
  scale_fill_viridis_c() +
  theme_minimal()

  ggplot(data, aes(mod_work, vig_work)) +
  geom_smooth(aes(color = race), se = FALSE, method = "lm") +
  facet_wrap(~data_cycle) +
  scale_color_viridis_d() +
  coord_flip() +
  theme_minimal()
```

###Feedback
You do a great job of packaging the plots together in order to tell a comprehensive story. You may want to try ggarrage() here also, but I will leave that call to you. I went ahead and used theme_minimal() again and also tried ColorBrewer's Dark2 theme on the second plot. One of the issues that I am having on my own plots with viridis is that it plots qualitative data in a gradient so that it does not actually look qualitative. I am sure Daniel has an easy fix, but one way is to go with another palette maker. 
